Refactoring of Software Using Relational Database

### 2013-02-07 doc/related_work.tex
finished OODB and RDB databases

\subsection{Database Evolution}
\subsubsection{Evolution in Objec-Oriented Databases}
The problem of database schema evolution is well known issue which is discussed for a long time. There is proposed an approach for  schemes of object-oriented databases in \cite{Banerjee:SemanticsOfSchemaEvolutionInOODB} which based on a set of schema invariants (representing the properties of the schema) and a set of rules for invariant preserving schema changes. The impact of evolution on instances is considered too. The approach is presented informally, while our approach is based on set of formal definitions.

>>>>

\subsection{Database Evolution}
The problem of database schema evolution is well known issue which is discussed for a long time, nevertheless the term "database refactoring" is quite new. This section provides an overview of significant related work in area of database evolution and refactoring.

\subsubsection{Evolution of Relational Database}
\paragraph{Informal definitions} The taxonomy of relational database evolution based on the entity-relationship model  is proposed in \cite{Roddick:TaxonomyOnERM}. The evolution is described as a change in entity-relationship model and change in relational database.  The semantics change patterns in context of conceptual schema is described in \cite{Wedemeijer:SemanticChangePatternsInConceptualSchema}, although its impact on database schema or data is not described. The main cases of data evolution are defined in both publications, however the description is informal. The extensive set of possible database refactorings is provided in \cite{Ambler:DbRefactoringBook}, where both schema and data evolution is discussed. The refactorings are intended to be used by database administrators, thus it assume database-first approach to evolution whereas this proposal is application-first.

\paragraph{Formal framework} A general formal framework for database evolution is defined in \cite{McBrien:formal-framework-transformation}. The framework is based on a set of basic graph transformations which are then extended to transformations of the entity-relationship model. The framework and the defined transformations can be used for comparison with our framework, where our framework consider stored data. The contribution of the formal framework is definition of equivalent structures in schemas.

\paragraph{Tools for Database Evolution} The MeDEA project \cite{Meda:Main}, \cite{Meda:MDDViewBasedEv} offers a tool for evolution of both database schema and stored data based on model-driven approach. Both MeDEA and proposed framework allows automatic database evolution, the difference is in different target audience.  The framework is aimed to be used by database administrators, therefore it is aimed on database structures (such as views) which are not considered in our approach as it is aimed to be used by developers. Proposed framework can use MeDEA to enrich its ability on database model, whereas MeDEA can use the code-centric point of view to design developer friendly transformations.

The PRISM is a research project for data management under schema evolution \cite{PRISM} in contrast with our proposal and with MeDEA it uses it extends the SQL command set by so called schema modification operators which implements the schema evolution. The project is aimed to  be used by database administrators (as well as MeDEA). The PRISM's set of modification operators could be used for database transformation in our proposal, however we prefer to work with a standard SQL set of commands.

\paragraph{ORM Frameworks} here are many object-relational mapping frameworks available for developers, and some of them provide tools for database migration. Hibernate \cite{Hibernate} is one of the most popular ORM frameworks in the Java community.  It provides customizable ORM for a wide range of databases, however it does not support complex database migrations.  It is capable only of creating a new table or adding a new column, hence it is not possible to, for example, drop a table or copy values from one column to another. Active Record \cite{Active_Record} is an ORM framework in the Ruby on Rails environment. Since its first version it has contained support for database evolution according to the create-update-delete principle, in the form of so-called migrations \cite{Rails:Migrations} which can be extended by adding user SQL commands. Entity Framework \cite{Entity_Framework} is Microsoft's ORM solution for the .NET platform. Its capabilities of data evolution support are similar to those of Active Record.

Each of presented ORM frameworks provides support for creating data evolution, as they are all capable of adding a new class as an entity and creating a corresponding table in the database. On the other hand, only two of them (Active Record and Entity Framework) are capable of updating and deleting evolution. In the case of evolutions which need to manipulate data, e.g. moving the property, none of the frameworks mentioned provides a built-in solution. Active Record and Entity Framework allow developers to describe the data move manually. In contrast to these ORMFs, our proposal  provides a higher level of user comfort because of its capability to handle complex data evolutions.

\subsubsection{Evolution of Object-Oriented Databases}
The issue of database evolution is discussed not only for relational databases, but for object-oriented databases too. There is proposed an approach for  schemes of object-oriented databases in \cite{Banerjee:SemanticsOfSchemaEvolutionInOODB} which based on a set of schema invariants (representing the properties of the schema) and a set of rules for invariant preserving schema changes. The impact of evolution on instances is considered too. The approach is presented informally, while our approach is based on set of formal definitions.
